---
title: "Data 607 Assignment 2"
author: "Dennis Pong"
date: "2/9/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

### Movie Rating Analysis using SQL and R

Step 0: Loading needed libraries 

Github location - 
https://github.com/metis-macys-66898/data_607_sp2020/blob/master/setup_script.sql


```{r}
library(tidyverse)
# install.packages("RPostgreSQL")
# install.packages("keyring")
# install.packages("chron")
require("RPostgreSQL")
require("keyring")

library(DBI)
require(chron)

library(tibble)
#library(reshape2)
library(ggplot2)
```


### Establishing Database Connection (RDMBS: Postgres)

* Establish the Postgres database connection with PostgreSQL Driver and proper connection strings:

+ user id (*user*)
+ password 
+ host
+ port number (*port*) 
+ database name (*dbname*)

At the end, we will do a check to make sure the database objects (ratings, movies, reviews) exists. 

STEP 1: Load Movie_Rankings.sql into R dataframe


```{r Step1}


# create a keyring 

keyring::key_set(service = "postgres", 
                 username = "postgres")

# keyring::key_list("postgres")[1,2]

# create a connection
# save the password that we can "hide" it as best as we can by collapsing it
# delete the following line if keyring is working
# pw <- {
#  "pgAdmin4"
#}
 

# loads the PostgreSQL driver
drv <- dbDriver("PostgreSQL")
# creates a connection to the postgres database
# note that "con" will be used later in each connection to the database
conn <- dbConnect(drv, dbname = "postgres",
                 host = "localhost", port = 5433,
                 user = keyring::key_list("postgres")[1,2], 
                 password = keyring::key_get("postgres", "postgres")
                 #password = pw
                )
# rm(pw) # removes the password

dbListTables(conn)


# check for the movies_survey_data
#dbExistsTable(conn, "ratings")
#dbExistsTable(conn, "movies")
#dbExistsTable(conn, "reviews")

# TRUE



```

Step 2: load the query results into data frames and join the Movie Data and Rating Data

#### Movie Table

```{r movies, echo=FALSE}
df_movies <- suppressWarnings(dbSendQuery(conn, "SELECT * from movies"))
movie_df <- fetch (df_movies)
movie_df

```

#### Reviewer Table 

```{r reviews}
df_reviewers <- suppressWarnings(dbSendQuery(conn, "SELECT * from reviews"))
reviewer_df <- fetch (df_reviewers)
reviewer_df
```



#### Rating Table
```{r ratings}
df_ratings <- suppressWarnings(dbSendQuery(conn, "SELECT * from ratings"))
rating_df <- fetch(df_ratings)
head(rating_df,10)


```

#### Joining Movie Data and Rating Data

```{r}
movie_ratings_reviewers <- left_join(movie_df, rating_df, by.x = "movie_id", by.y = "movie_id")
```

check the combined dataframe and view the data

```{r}
movie_ratings_reviewers
```


STEP 3: Handling Null Values

examine the severity of missing values 
```{r missing-value-check}
movie_ratings_reviewers %>% filter(is.na(rating))
```

Imputting missing rating values with mean rating.

```{r}

movie_reviewer_rating_cleaned_df <- movie_ratings_reviewers %>% mutate (
                                                                      rating = chron::times(rating), 
                                                                      rating = if_else(is.na(rating), mean(rating), rating)
                                                                        )

movie_reviewer_rating_cleaned_df

movie_reviewer_rating_cleaned_df %>% filter(is.na(rating))


```

> results should show there is no more NA


STEP 4: Summarize the data 

```{r}

movie_ratings <- movie_reviewer_rating_cleaned_df %>% group_by(movie_name) %>% summarize(movie_rating = mean(rating))  %>% arrange(desc(movie_rating))

movie_ratings
    

```

STEP 5: Graph Data Distribution

```{r}



movie_reviewer_rating_cleaned_df_2 <- left_join (movie_reviewer_rating_cleaned_df, reviewer_df, by.x="reviewer_id", by.y = "reviewer_id"  )
movie_reviewer_rating_cleaned_df_2

final_df <- movie_reviewer_rating_cleaned_df_2 %>% select(movie_name, reviewer_name, rating)

# plotting a boxplot of ratings by movie name
# Boxplot (1)
ggplot(final_df, aes (x = movie_name, y = rating )) + geom_boxplot() 

# plotting a boxplot of ratings by reviewer name
# Boxplot (2)
ggplot(final_df, aes (x = reviewer_name, y = rating )) + geom_boxplot() 

```


Step 6: CLOSE the db connection

```{r}
# Closing all active connections is important
dbDisconnect(conn)
#dbListConnections( dbDriver( drv = "PostgreSQL"))
lapply( dbListConnections( dbDriver( drv = "PostgreSQL")), dbDisconnect)
```


### Conclusion 

After handling the missing values present in the movie_ratings_reviewers dataset, the empirical results show that **Your Name** has the highest rating and **Rogue One: A Star Wars Story**  finished last. The boxplot (1) shows that Rogue One does have a wider spread in ratings than any other movies. 
