---
title: "Week 3 assignment - DATA 607"
author: "Dennis Pong"
date: "2/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#1. Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"


```{r}
library(httr)

url  <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv"
majors <- read.csv(paste0(url), header = TRUE)
grep(pattern = 'STATISTICS|DATA', majors$Major, value = TRUE, ignore.case = TRUE)

```


#2 Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

```{r}

library(stringr)

startstr <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"'

dbl_quote = '"'
# Use function gregexpr to extract the substring = pattern
double_quotes_positions <- gregexpr(pattern =  dbl_quote, text = startstr)
# check double_quotes_positions[[1]][1]
double_quotes_positions[[1]][1]
# print head of double_quotes positions
head(double_quotes_positions[[1]])

# store all the double_quotes positions into the vector dq_pos
dq_pos <- vector()
i <- 1
while (!is.na(double_quotes_positions[[1]][i])){
                              dq_pos[i] <- double_quotes_positions[[1]][i]
                              i <- i+1
}

no_of_words <- length(dq_pos)/2
desired_output <- vector (length=no_of_words)
#print(desired_output)
# checking the length of desired_output 
length(desired_output)

for (i in 1:no_of_words) {
desired_output[i] <- substring(startstr,double_quotes_positions[[1]][2*i-1]+1,double_quotes_positions[[1]][2*i]-1)
i <- i+1
}

# set the optional character to be \", \"
# Use writeLines to complete the deal
end_result <- paste0("c(\"", paste0(desired_output, collapse = "\", \""), "\")")


writeLines(end_result)
```





The two exercises below are taken from R for Data Science, 14.3.5.1 in the on-line version:

#3 Describe, in words, what these expressions will match:

* `(.)\1\1 `- This regular expression matches a string containing the same three consecutive characters. It would match, for example, “abbbc” but not “abc”.  
* `"(.)(.)\\2\\1"`  - This is a string representing a regular expression that matches a pair of any characters followed by the reverse order of the same pair. So, “abbbba” and “abba” are both viable matches.  
* `(..)\1 `- This regular expression matches any two characters followed by the same sequence of two characters.  
* `"(.).\\1.\\1"`  - five characters where the first, third and fifth are the same and the second and fourth can be anything. Possible matches could be “abaaa”, “dedad”.  
* `"(.)(.)(.).*\\3\\2\\1"` - 6 or more carachters where the first three characters are the same as the last three in reverse order. It could be of length 7 or 6. If it's of length 7, the fourth character 
can be anything. Possible matches could be something like “abcdcba” or “abccba”  


#4 Construct regular expressions to match words that:

* Start and end with the same character. - `"^(.).*\1$"`  

* Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.) - `".*(.?...?).*\\1"  `  

* Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.) - `"(.?[A-Za-z].?){3,}"  `  


```{r this-matching-program-doesn\'t-work-in-some-cases, eval=FALSE, include=FALSE}
regex_expr = '"^(.).*\1$"'

test_cases = c('"11"', '"231"', '"232"', '231', '"a0a"', '"aa20"', '"a"')

str_view_all(test_cases, regex_expr)
```


