---
title: "Week 9 Assignment - Web APIs"
author: "Dennis Pong"
date: "3/29/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Introduction

This RMD connects to NYTimes' Most Popular API, reads in JSON data, and transforms it into a R Dataframe.

**Most Popular API**  
  
> Provides services for getting the most popular articles on NYTimes.com based on emails, shares, or views.

### Setup

```{r}
library(jsonlite)
library(dplyr)
```

#### Use an alternative but a more consistent mapping to rjson and RJSONIO to convert JSON strings to R objects. Turned on the flatten option to automatically flatten nested data frames into a single non-nested data frame. Skipped the step to use flattern function later on.

#### Top 10 most emailed articles in the past 7 days (published date AND URL)

```{r data-retrieval_(i)}

most_emailed_articles <- fromJSON("https://api.nytimes.com/svc/mostpopular/v2/emailed/7.json?api-key=BLtybA60SxbT2XcgepyXjVVGOQUCn6iR", flatten = TRUE) %>% data.frame()

# str(most_emailed_articles)
# dim(most_emailed_articles)



most_emailed_articles %>% select (results.published_date, results.url) %>% head(10)
```

#### Top 10 most shared articles on facebook in the past 1 day (published date AND URL)


```{r data-retrieval-(ii)}
# Most shared articles on Facebook for the last day

most_shared_articles_fb <- fromJSON("https://api.nytimes.com/svc/mostpopular/v2/shared/1/facebook.json?api-key=BLtybA60SxbT2XcgepyXjVVGOQUCn6iR", flatten = TRUE) %>% data.frame()

# str(most_shared_articles_fb)
# dim(most_shared_articles_fb)

most_shared_articles_fb %>% select (results.published_date, results.url) %>% head(10)

```

#### Top 10 most viewed articles in the past 30 days (published date AND URL)


```{r data_retrival_(iii)}

most_viewed_articles <- fromJSON("https://api.nytimes.com/svc/mostpopular/v2/viewed/30.json?api-key=BLtybA60SxbT2XcgepyXjVVGOQUCn6iR", flatten = TRUE) %>% data.frame()

# str(most_viewed_articles)
# dim(most_viewed_articles)

most_viewed_articles %>% select (results.published_date, results.url) %>% head(10)



```

### Final Remarks

#### Figured out that this API only takes 3 different periods of time, namely, 1, 7, 30 days in the past. 
