---
title: "Dennis Pong - Project1_D607"
author: "Dennis Pong"
date: "2/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r laoding-libraries}
library(dplyr)
library(stringr)
library(tidyr)

```



## Loading of the tournamentinfo text file


```{r loading-tourney-data}

data<-readLines('https://raw.githubusercontent.com/metis-macys-66898/data_607_sp2020/master/tournamentinfo.txt', warn = 'F')


data<-data[data != strrep("-", 90-1)]

#head(data)
#length(data)

data<-data[3:130]
head(data)

```

## The way that we're set up is we started to import the text file using readLines as a character vector. Then I'm going to use regexp to parse it into individual character vectors which are the variables that we can use to construct a final dataframe. 


```{r}


# Capturing names like CAMERON WILLIAM MC LEMAN and SEAN M MC CORMICK
# assuming the name could max out the field length allowed, I do not expect a space to be matched in the end for 2-word, 3-word, 4-word names. 

name_regex<-"[0-9] \\| [a-zA-Z]+\\s[a-zA-Z]+\\s?([a-zA-Z]+\\s?)*[a-zA-Z]*"
# removing the leading and trailing space along with the leading non-named related string that I wanted to remove
name<- trimws(str_remove_all(unlist(str_extract_all(data, name_regex)), "\\d\\s\\|"))
name



state_regex<-"\\s{2}[A-Z]{2}"
states<-unlist(str_extract_all(data, state_regex))
states


points_regex<-"\\d\\.\\d"
points<-unlist(str_extract_all(data, points_regex))
points



pre_rating_regex<-"R:\\s+[0-9]{3,}"
pre_rating<-str_remove_all(unlist(str_extract_all(data, pre_rating_regex)), "R:\\s")
pre_rating


uscf_regex<-"[0-9]{8}"
USCF_ID<-unlist(str_extract_all(data, uscf_regex))
USCF_ID
# length(uscf_id)


# this will match anything that is W, L, or D, but only exclude other types from the two ends of the bar (|)
games_regex<-"\\|[W|L|D].+[0-9]{1,2}\\|"
# Having this games_regex_removed ensures we do not take in any B, H, or U
games_regex_removed <- "\\|[^(W|L|D)]\\s{4}"
games<- unlist(str_extract_all(data, games_regex))
games<-str_remove_all(games, games_regex_removed)
# str_extract_all with simplify defaulted to F gives us a list of character vectors, which can then be unlisted to collapse to 64 elements
games<-strsplit(unlist(str_extract_all(data, games_regex)), "\\|")
# after splitting on string literal bar (|), we should end with 64 elements in the list
length(games)

# creating the list of games for all 64 players
for (i in 1:length(games)){
  games[[i]]<-games[[i]][2:length(games[[i]])]

}




```


## To answer the question for getting the average of the Pre Chess Rating of Opponents for players 1 and 37 in the Intro YouTube video, we create a function to get the opponent_scores needed to create the data frame for the final output. Need to use extra caution here as most of the variables are defaulted to factors. I need to explicitly change it back to chracter, numeric, or integers. Note that while writeLines has its merits, I do think read.csv is another standard approach where you can set stringsAsFactors = F to take care of the Factors issues. 


```{r attempt1}

opponent_score <- function(game_num) {
  game_outcome <- rep(NA,64)
  game_opponent_rank <- rep(NA, 64)
  
for (i in 1:length(games)){
    # this pulls out the 1-character W, L, D 
    game_outcome[i]<-str_extract(games[[i]][game_num],"[A-Za-z]{1}") 
    # This gives me the opponent rank of a given game i
    game_opponent_rank[i]<-str_extract(games[[i]][game_num],"[0-9]{1,2}")
    
}
  result <- list(game_outcome, game_opponent_rank)
  return (result)
}

#opponent_score(1)[[2]]

df <- data.frame(USCF_ID, 
                 name, 
                 states,
                 points, 
                 pre_rating, 
                 opponent_score(1)[[2]], 
                 opponent_score(2)[[2]],
                 opponent_score(3)[[2]], 
                 opponent_score(4)[[2]],
                 opponent_score(5)[[2]], 
                 opponent_score(6)[[2]],
                 opponent_score(7)[[2]]
                )
#df
#head(df,2)
#Renaming opponent_score for the 7 columns 
names(df) <- c("USCF_ID", "name", "states", "points", "pre_rating", "opp_rating_1", "opp_rating_2", "opp_rating_3", "opp_rating_4", "opp_rating_5", "opp_rating_6", "opp_rating_7")
names(df)

#str(df)
# points scored needed to be in numeric 
df$points<-as.numeric(as.character(df$points))
# USCF_ID along with all the opponent scores, and pre_Ratings are supposed to be all integers
cols <- c(names(df[, 5:12]), "USCF_ID")
df[,cols] <- apply(df[,cols], 2, function(x) as.integer(as.character(x)))

#str(df)
#df

char_cols <- c("name", "states")
df[,char_cols] <- apply(df[,char_cols], 2, function(x) as.character(x))

#col_names <- names(df[,6:12])
opp_ratings<- c('opp_rating_1', 'opp_rating_2', 'opp_rating_3', 'opp_rating_4', 'opp_rating_5', 'opp_rating_6', 'opp_rating_7')

populate_opp_scores <- function (df, cols){
for (j in opp_ratings) {
  for (i in 1:nrow(df)){
    df[[j]][i]<- df$pre_rating[df[[j]][i]]
  }
  
}
  return (df)
}



df<- populate_opp_scores(df, opp_ratings)
#df
#df[37,]

# calculating row-level means using rowMeans
df$average_opponent_rating <- rowMeans(df[,opp_ratings], na.rm = T)

#df
# Do not need to use group_by if the dataset is large. 
final_df <- select( df, -USCF_ID, -all_of(opp_ratings) ) %>% group_by(name, states, points, pre_rating) %>%  arrange(desc(average_opponent_rating))

final_df



#head(final_df) 


```

## Writing out the file to csv to complete the task 

```{r write-out-to-csv}


write.csv(final_df, "project_1_DataFrame_DP.csv")
cur_dir<-getwd()
list.files(cur_dir)

```
## Bonus Track:
### Furthermore, I'd like to examine the tournament results by comparing the average_opponent_rating and pre_rating (pre-tournament rating) and see if there is an origin that has a bigger difference than the others. It turns out there are only 3 states, missouri (MI), Ohio (OH), and Ontario, Canada (ON) that we have data for. OH has only 1 count of non-NA. I'm going to exclude it from comparison. It turns out that ON has a higher net tournament results, defined by pre-tournament rating minus average opponent rating. 


```{r}

final_df2<-final_df
final_df2$tournament_results <- final_df$pre_rating - final_df$average_opponent_rating
#final_df2


boxplot( tournament_results ~ states, data = final_df2)

#str(final_df2)
final_df2 %>%  group_by(states) %>% summarize(counts  = n(), na.rm = T)
```

